---
title: "Assignment 4"
author: "Gustav Oosthuizen - OSTGUS001"
date: '`r format(Sys.Date(), "%d-%B-%Y")`'
output:
  pdf_document:
    highlight: pygments
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_caption: true
citation_package: "biblatex"
bibliography: "Ass_4.bib"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\pagebreak

```{r Installing packages, include=FALSE}

# Steps:

# 1. Installing packages used for analysis:

# 1.1 General use

#install.packages("tidyverse")
#install.packages("remotes")
#install.packages("rgdal")
#install.packages("devtools")
#install.packages("RColorBrewer")
#install.packages("ggedit")
#install.packages("caTools")
#install.packages("plotly")
#install.packages("gridExtra")
#install.packages("egg")
install.packages("cowplot")

library(cowplot)
library(tidyverse)
library(remotes)
library(rgdal)
library(devtools)
library(RColorBrewer)
library(ggedit)
library(caTools)
library(plotly)
library(gridExtra)
library(egg)

# 1.2 Mapping data and working with spatial objects

#install.packages("spdplyr")
#install.packages("sf")
#install.packages("countrycode")
#install.packages("ggmap")
#install.packages("rnaturalearth")
#install.packages("rnaturalearthdata")
#install.packages("mapdata")
#intall,packages("leaflet")

library(spdplyr)
library(sf)
library(countrycode)
library(ggmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(leaflet)
library(mapdata)

# 1.3 Creating animations

#install.packages("gganimate")
#install.packages("gifski")


library(gifski)
library(gganimate)

# 1.4 Creating Shiny dashboard


#install.packages("shiny")
#install.packages("shinydashboard")
#install.packages("DT")
#install.packages("shinyWidgets")
#install.packages("shinybusy")



library(shinybusy)
library(shinyWidgets)
library(shinydashboard)
library(DT)
library(shiny)

```


\section{Data wrangling} \label{DW section}

The data wrangling consist of importing, tidying and transforming the data set [@wickham2016r].

\subsection{Imporing data} 

As mentioned in section \ref{DW section} the first step of data wrangling is importing the data.

***** START CODE - for reproduceability *****

```{r Loading covid 19 data}

# Note: The original data file will be attached for reproducability.

# Steps:

# 1. Loading data 


# 1.1 Covid 19 data up to the 23/03/2020 (Most code written for this data set)

# ncov_new <- read.csv(file = "time_series_19_covid_combined_23_03_2020.csv", stringsAsFactors = FALSE)


# 1.2 Covid 19 data up to the 15/04/2020 (Test data set)

ncov_new <- read.csv(file = "time_series_19_covid_combined_15_04_2020.csv", stringsAsFactors = FALSE)

```


\subsection{Tiding and transforming the data}

The second and thrid steps of data wrangling is tiding and transforming the data. The variables of the "ncovr_new" data set will be investigated and the nesseccary changes/transformations will be made to get the data in a tidy format.   


```{r Investigating variables and treating missing values in data sets}

# Steps:

# 1. Investigating type and structre of data

glimpse(ncov_new)

str(ncov_new)

# Remark: The "Date" variable should of type date not character:


# 2. Checking for NA values in data set

sum(is.na(ncov_new$Province.State))
sum(is.na(ncov_new$Country.Region))
sum(is.na(ncov_new$Lat))
sum(is.na(ncov_new$Long))
sum(is.na(ncov_new$Date))
sum(is.na(ncov_new$Confirmed))
sum(is.na(ncov_new$Recovered))
sum(is.na(ncov_new$Deaths))

# Remark: The "Confirmed" and Deaths variables seem to have missing values

# 2.1 Looking at missing value instances

ncov_new %>% 
  filter(is.na(Confirmed))

# Remark: It seems to be associated with a sertain location in Canada. The "Recovered" variable is populated but not the "Deaths" or "Confirmed" variables. It would be best to totally ignore these instances since they offer no value. All three variables need to be known for data to be useful.

# 3. Remove instances with NA values

ncov_new <- na.omit(ncov_new)


```


```{r Additions and transformations to data set}

# Steps:

# 1. Changing the format of the "Date" variable

ncov_new$Date <- as.Date(ncov_new$Date,)

# 2. Adding a variable "Infected" which is the number of accounts of people that are still undergoing treatment.

ncov_new <- ncov_new %>% 
  mutate(Infected = Confirmed - Recovered - Deaths)


# 3. Renaming some of the variables to make it more intuitive

ncov_new <- ncov_new %>% 
  rename(Cum_Confirmed = Confirmed,
         Province_State = Province.State,
         Country_Region = Country.Region)

```


```{r Loading of spatial data}

# Note: The following spatial data was obtained by installing the naturalearth packages.

# Steps:

# 1. Loading of the world data (spatial object) 

World <- ne_countries(scale = "medium", returnclass = "sf")
  
```


```{r Making neccessary changes to accomplish merging of ncov_new and World data frames}

# Note: The "Country_Region" variable will be used to join the spatial data frame ("World") and the ncov_new data frame. Therfore, a geometry variable will be added to the ncov_new data frame.  

# Steps:

# 1. Testing to see which country names do NOT coincide between the two data frames.

# 1.1 Creating a data frame (World_df) that list all the countries of the natural earth data.The "pop_est" variable will be used later on for testing purpose.

World_df <- data.frame( Country_Region = World$name_long,
                        Pop_Estimate = World$pop_est)


# 1.2 Joining the ncov_new data frame with the World_df date frame to investigate cases where "pop_est" is NA (i.e to investigate where there is a mis match in country names)

Country_Test_df <- left_join(ncov_new, World_df, by = "Country_Region" )


# Remark: There are NA's present. These country names that cause this will be investigated.

# 1.3 Investigation of the countries that present NA values

 Mismatch_Countries <- Country_Test_df %>%  
   filter(is.na(Pop_Estimate)) %>%
   group_by(Country_Region) %>%
   summarise(count = n())

 # Remark: Luckily, there is only 20 counties that does not coincide. These countries will be manually renamed so that they can be merged successfully.  
 
# 1.4  Change the following names of the ncov_new data set. Format: "old name" -> "new name"

# 1. Brunei -> Brunei Darussalam 
# 2. Congo (Brazzaville) -> Republic of the Congo
# 3. Congo (Kinshasa) -> Democratic Republic of the Congo
# 4. Cote d'Ivoire -> Côte d'Ivoire 
# 5. Diamond Princess -> Japan 
# 6. Czechia -> Czech Republic 
# 7. Eswatini -> eSwati
# 8. Gambia, The ->The Gambia
# 9. Holy See -> Vatican 
# 10. Korea, South -> Republic of Korea 
# 11. North Macedonia -> Macedonia 
# 12. Russia -> Russian Federation 
# 13. Taiwan* -> Taiwan 
# 14. US -> United States 
# 15. Cabo Verde -> Republic of Cabo Verde 
# 16. Burma -> Myanmar 
# 17. Laos -> Lao PDR
# 18. MS Zaandam -> Bahamas
# 19. Sao Tome and Principe -> 	São Tomé and Principe
# 20. West Bank and Gaza -> Palestine

ncov_new$Country_Region[ncov_new$Country_Region == "Brunei"] <- "Brunei Darussalam"

ncov_new$Country_Region[ncov_new$Country_Region == "Congo (Brazzaville)"] <- "Republic of Congo"

ncov_new$Country_Region[ncov_new$Country_Region == "Congo (Kinshasa)"] <- "Democratic Republic of the Congo"

ncov_new$Country_Region[ncov_new$Country_Region == "Cote d'Ivoire"] <- "Côte d'Ivoire"

ncov_new$Country_Region[ncov_new$Country_Region == "Diamond Princess"] <- "Japan"

ncov_new$Country_Region[ncov_new$Country_Region == "Czechia"] <- "Czech Republic"

ncov_new$Country_Region[ncov_new$Country_Region == "Eswatini"] <- "Swaziland"

ncov_new$Country_Region[ncov_new$Country_Region == "Gambia"] <- "The Gambia"

ncov_new$Country_Region[ncov_new$Country_Region == "Holy See"] <- "Vatican"

ncov_new$Country_Region[ncov_new$Country_Region == "Korea, South"] <- "Republic of Korea"

ncov_new$Country_Region[ncov_new$Country_Region == "Russia"] <- "Russian Federation"

ncov_new$Country_Region[ncov_new$Country_Region == "US"] <- "United States"

ncov_new$Country_Region[ncov_new$Country_Region == "Cabo Verde"] <- "Cape Verde"

ncov_new$Country_Region[ncov_new$Country_Region == "North Macedonia"] <- "Macedonia"

ncov_new$Country_Region[ncov_new$Country_Region == "Taiwan*"] <- "Taiwan" 

ncov_new$Country_Region[ncov_new$Country_Region == "Burma"] <- "Myanmar"

ncov_new$Country_Region[ncov_new$Country_Region == "Laos"] <- "Lao PDR" 

ncov_new$Country_Region[ncov_new$Country_Region == "MS Zaandam"] <- "Bahamas" 

ncov_new$Country_Region[ncov_new$Country_Region == "West Bank and Gaza"] <- "Palestine" 

ncov_new$Country_Region[ncov_new$Country_Region == "Sao Tome and Principe"] <- "São Tomé and Principe" 


# 1.5 Re-joining the Country_Test_df with changed country names

Country_Test_df <- left_join(ncov_new, World_df, by = "Country_Region")


# Remark: All the country names of the ncov_new data set and the spatial data set will coincide. The merging of the spatial features can continue.


# 2. Creating Spatial data frame that only contains the important features from the world data set and storing in "World_Merge_df". This spatial data frame will be used in the merging operation in the next code chunk.

World_Merge_df <- World %>% 
  select(name_long,continent) %>%
  rename(Country_Region  = name_long)

```


```{r Merging operation of spatial data}

# Note: The following code chunk deals with creating a data frame that contains the nessecceray data that will be used to create plots of a world map.


# Steps:

# 1. Creating a data frame that summarises the data for each country per day. 

ncov_new_FINAL_data <-  ncov_new %>% 
  group_by(Date, Country_Region) %>% 
  summarise(Cum_Confirmed_Country = sum(Cum_Confirmed),
            Recovered_Country = sum(Recovered),
            Deaths_Country = sum(Deaths),
            Infected_Country = sum(Infected)) %>% 
  as.data.frame()

# Remark: The Province_State varibale was ignored in the analysis since most of the values were missing.

# 2. Merge data frame (ncov_new_FINAL_data) with Spatial data frame (World_Merde_df)

ncov_new_Spatial <- left_join(ncov_new_FINAL_data, World_Merge_df , by = "Country_Region")

# 3. Classifying the geometry feature in data frame.The data frame will then have a class "data frame" and "sf" instead of just a "data frame"

st_geometry(ncov_new_Spatial) <- ncov_new_Spatial$geometry

```


\section{Visualisations}

\subsection{Part 1 - Animation development}


```{r  Creating an animation of the world map}

# Note: The following code blocks will be associated with creating an animations to give insight on the spread of covid 19 gloabally. Creating a single animation of the world map, showing the cumalative cases of covid 19 is insightful, but there is a lack in detail. Therefore it was decided to make multiple animations (gifs) and zooming into areas of intrest to observe more detail.  

# Steps: 

# 1. Specifying the position of the labels for each country (counting the number of cases as the date progresses)

World_points<- st_centroid(ncov_new_Spatial)
World_points <- cbind(ncov_new_Spatial,
                      st_coordinates(st_centroid(ncov_new_Spatial$geometry)))


# 2. Plotting world map using ggplot()

 World_Map <- ggplot() +
   geom_sf(data = World, fill = "gray48") + 
   geom_sf(data = ncov_new_Spatial, aes(frame = Date, fill = Cum_Confirmed_Country)) +
   scale_fill_gradient(low="white", high="red", trans = "log10") + 
   labs(title = "Cumalative confirmed cases of covid 19 - World Map", subtitle = "Date:{current_frame}", x = "Latitude",y = "Longitude", fill = "Count") + 
   theme(plot.title = element_text(size = 14, face = "bold"), 
         panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))
   

# 3. Creating a gganim object
 
World_Map_anim <-  World_Map + transition_manual(Date)

# 4. Creating gif

World_Map_Gif <- animate(World_Map_anim, fps = 5, width = 950, height = 850)

# 5. Saving the gif created (Maybe should not insert)

anim_save("1.World_Map.gif", animation = World_Map_Gif, path = "C:/Users/Gustav/Desktop/MSc Data Science 1st year/Exploratory data analysis/Assignments/Assignment4" )
```


```{r Creating an animation for Africa}

# Steps:

# 1. Filtering the spatial data frame to only retrieve information about Africa

ncov_new_Spatial_Afr <- ncov_new_Spatial %>%
  filter(continent == "Africa")

# 2. Specifying the position of the labels for each country in Africa (counting the number of cases as the date progresses)

World_Points_Afr <- st_centroid(ncov_new_Spatial_Afr)
World_Points_Afr <- cbind(ncov_new_Spatial_Afr,
                          st_coordinates(st_centroid(ncov_new_Spatial_Afr$geometry)))


# 3. Getting data for base map (Africa countries)

Africa_data <- World %>% 
  filter(continent == "Africa")

# 4. Plotting Africa map using ggplot()

Map_Afr <- ggplot() + 
  geom_sf(data = Africa_data, fill = "gray48" ) +
  geom_sf(data = ncov_new_Spatial_Afr, aes(frame = Date, fill = Cum_Confirmed_Country)) +
  scale_fill_gradient(low="white", high="red", trans = "log10") +
  geom_text(data = World_Points_Afr, aes(x = X, y = Y, label = Cum_Confirmed_Country), check_overlap = FALSE) + 
  labs(title = "Cumalative confirmed cases of covid 19 - Africa", subtitle = "Date: {current_frame}", fill = "Count", x = "Latitude", y = "Longitude" ) + 
  theme(title = element_text(size = 14, face = "bold"), panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))
 
# 5. Creating a gganim object
 
Map_Afr_Anim <-  Map_Afr + transition_manual(Date)

# 6. Creating gif

Afr_Gif <-  animate(Map_Afr_Anim, fps = 5, width = 950, height = 750)

# 7. Saving the gif created (Maybe should not insert)

anim_save("2.Afr.gif", animation = Afr_Gif, path = "C:/Users/Gustav/Desktop/MSc Data Science 1st year/Exploratory data analysis/Assignments/Assignment4" )

```


```{r Creating an animation for Europe}

# Steps:

# 1. Filtering the spatial data frame to only retrieve information about Europe

ncov_new_Spatial_Eur <- ncov_new_Spatial %>% 
  filter(continent == "Europe")

# 2. Specifying the position of the labels for each country in Africa (counting the number of cases as the date progresses)

World_Points_Eur <- st_centroid(ncov_new_Spatial_Eur)
World_Points_Eur <- cbind(ncov_new_Spatial_Eur,
                          st_coordinates(st_centroid(ncov_new_Spatial_Eur$geometry)))

# 3. Getting data for base map (Europe countries)

Europe_data <- World %>% 
  filter(continent == "Europe")

# 4. Plotting Europe map using ggplot() 

Map_Eur <- ggplot() + 
  geom_sf(data = Europe_data, fill = "gray48") +
  geom_sf(data = ncov_new_Spatial_Eur, aes(frame = Date, fill = Cum_Confirmed_Country)) +
  coord_sf(xlim = c(-40, 100), ylim = c(30,75), expand = FALSE) +
  scale_fill_gradient(low="white", high="red", trans = "log10") +
  geom_text(data = World_Points_Eur, aes(x = X, y = Y, label = Cum_Confirmed_Country), check_overlap = FALSE) + 
  labs(title = "Cumalative confirmed cases of covid 19 - Europe", subtitle = "Date: {current_frame}", fill = "Count", x = "Latitude", y = "Longitude" ) + 
  theme(title = element_text(size = 14, face = "bold"), panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))

# 5. Creating a gganim object
  
Map_Eur_Anim <-  Map_Eur  + 
  transition_manual(Date)

# 6. Creating gif 

Eur_Gif <-  animate(Map_Eur_Anim, fps = 5, width = 950, height = 750)

# 7. Saving the gif created (Maybe should not insert)

anim_save("3.Eur.gif", animation = Eur_Gif, path = "C:/Users/Gustav/Desktop/MSc Data Science 1st year/Exploratory data analysis/Assignments/Assignment4" )
```


```{r Creating an animation for Asia}

# Steps:

# 1. Filtering the spatial data frame to only retrieve information about Asia 

ncov_new_Spatial_Asia <- ncov_new_Spatial %>% 
  filter(continent == "Asia")

# 2. Specifying the position of the labels for each country in Asia (counting the number of cases as the date progresses) 

World_Points_Asia <- st_centroid(ncov_new_Spatial_Asia)
World_Points_Asia <- cbind(ncov_new_Spatial_Asia,
                           st_coordinates(st_centroid(ncov_new_Spatial_Asia$geometry)))

# 3. Getting data for base map (Asia countries)

Asia_data <- World %>% 
  filter(continent == "Asia")

# 4. Plotting Asia map using ggplot()  

Map_Asia <- ggplot() + 
  geom_sf(data = Asia_data, fill = "gray48") +
  geom_sf(data = ncov_new_Spatial_Asia, aes(frame = Date, fill = Cum_Confirmed_Country)) +
  scale_fill_gradient(low="white", high="red", trans = "log10") +
  geom_text(data = World_Points_Asia, aes(x = X, y = Y, label = Cum_Confirmed_Country), check_overlap = FALSE) + 
  labs(title = "Cumalative confirmed cases of covid 19 - Asia", subtitle = "Date: {current_frame}", fill = "Count", x = "Latitude", y = "Longitude" ) + 
  theme(title = element_text(size = 14, face = "bold"), panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))

# 5. Creating a gganim object

Map_Asia_Anim <-  Map_Asia  + 
  transition_manual(Date)

# 6. Creating gif

Asia_Gif <-  animate(Map_Asia_Anim, fps = 5, width = 950, height = 750)

# 7. Saving the gif created (Maybe should not insert)

anim_save("4.Asia.gif", animation = Asia_Gif, path = "C:/Users/Gustav/Desktop/MSc Data Science 1st year/Exploratory data analysis/Assignments/Assignment4" )

```


```{r Creating an animation for North America}

# Steps:

# 1. Filtering the spatial data frame to only retrieve information about North America 

ncov_new_Spatial_USA <- ncov_new_Spatial %>% 
  filter(continent == "North America")


# 2. Specifying the position of the labels for each country in North America (counting the number of cases as the date progresses) 

World_Points_USA <- st_centroid(ncov_new_Spatial_USA)
World_Points_USA <- cbind(ncov_new_Spatial_USA,
                          st_coordinates(st_centroid(ncov_new_Spatial_USA$geometry)))

# 3. Getting data for base map (North America countries)

USA_data <- World %>% 
  filter(continent == "North America")

# 4. Plotting North America map using ggplot() 

Map_USA <- ggplot() + 
  geom_sf(data = USA_data, fill = "gray48") +
  geom_sf(data = ncov_new_Spatial_USA, aes(frame = Date, fill = Cum_Confirmed_Country)) +
  coord_sf(xlim = c(-180,-10), ylim = c(5, 85), expand = FALSE) +
  scale_fill_gradient(low="white", high="red", trans = "log10") +
  geom_text(data = World_Points_USA, aes(x = X, y = Y, label = Cum_Confirmed_Country), check_overlap = FALSE) + 
  labs(title = "Cumalative confirmed cases of covid 19 - North America", subtitle = "Date: {current_frame}", fill = "Count", x = "Latitude", y = "Longitude" ) + 
  theme(title = element_text(size = 14, face = "bold"), panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))

# 5. Creating a gganim object 

Map_USA_Anim <-  Map_USA  + 
  transition_manual(Date)

# 6. Creating gif

USA_Gif <-  animate(Map_USA_Anim, fps = 5, width = 950, height = 750)

# 7. Saving the gif created (Maybe should not insert)

anim_save("5.USA.gif", animation = USA_Gif, path = "C:/Users/Gustav/Desktop/MSc Data Science 1st year/Exploratory data analysis/Assignments/Assignment4" )

```

```{r Creating an animation for Oceania}

# Steps:

# 1. Filtering the spatial data frame to only retrieve information about Oceania

ncov_new_Spatial_Oce <- ncov_new_Spatial %>% 
  filter(continent == "Oceania")

# 2. Specifying the position of the labels for each country in Oceania (counting the number of cases as the date progresses) 

World_Points_Oce <- st_centroid(ncov_new_Spatial_Oce)
World_Points_Oce <- cbind(ncov_new_Spatial_Oce,
                          st_coordinates(st_centroid(ncov_new_Spatial_Oce$geometry)))

# 3. Getting data for base map (Oceania countries)  

Oceania_data <- World %>% 
  filter(continent == "Oceania")

# 4. Plotting Oceania map using ggplot()  

Map_Oce <- ggplot() + 
  geom_sf(data = Oceania_data, fill = "gray48") +
  geom_sf(data = ncov_new_Spatial_Oce, aes(frame = Date, fill = Cum_Confirmed_Country)) +
  coord_sf(xlim = c(110,190), ylim = c(-60,20), expand = FALSE) +
  scale_fill_gradient(low="white", high="red", trans = "log10") +
  geom_text(data = World_Points_Oce, aes(x = X, y = Y, label = Cum_Confirmed_Country), check_overlap = FALSE, size = 6) + 
  labs(title = "Cumalative confirmed cases of covid 19 - Oceania", subtitle = "Date: {current_frame}", fill = "Count", x = "Latitude", y = "Longitude" ) + 
  theme(title = element_text(size = 14, face = "bold"), panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))

# 5. Creating a gganim object 
  
Map_Oce_Anim <-  Map_Oce  + 
  transition_manual(Date)

# 6. Creating gif

Oce_Gif <-  animate(Map_Oce_Anim, fps = 5, width = 950, height = 750)

# 7. Saving the gif created (Maybe should not insert)

anim_save("6.Oce.gif", animation = Oce_Gif, path = "C:/Users/Gustav/Desktop/MSc Data Science 1st year/Exploratory data analysis/Assignments/Assignment4" )

```

```{r Creating an animation for South America }

# Steps: 

# 1.Filtering the spatial data frame to only retrieve information about South America

ncov_new_Spatial_SAm <- ncov_new_Spatial %>% 
  filter(continent == "South America")

# 2. Specifying the position of the labels for each country in South America (counting the number of cases as the date progresses) 

World_Points_SAm <- st_centroid(ncov_new_Spatial_SAm)
World_Points_SAm <- cbind(ncov_new_Spatial_SAm,
                          st_coordinates(st_centroid(ncov_new_Spatial_SAm$geometry)))

# 3. Getting data for base map (South America countries)

SAm_data <- World %>% 
  filter(continent == "South America")

# 4. Plotting South America map using ggplot() 

Map_SAm <- ggplot() + 
  geom_sf(data = SAm_data, fill = "gray48") +
  geom_sf(data = ncov_new_Spatial_SAm, aes(frame = Date, fill = Cum_Confirmed_Country)) +
  coord_sf(xlim = c(-100,-30), ylim = c(-60,20), expand = FALSE) +
  scale_fill_gradient(low="white", high="red", trans = "log10") +
  geom_text(data = World_Points_SAm, aes(x = X, y = Y, label = Cum_Confirmed_Country), check_overlap = FALSE) + 
  labs(title = "Cumalative confirmed cases of covid 19 - South America", subtitle = "Date: {current_frame}", fill = "Count", x = "Latitude", y = "Longitude" ) + 
  theme(title = element_text(size = 14, face = "bold"), panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))
 
# 5. Creating a gganim object

Map_SAm_Anim <-  Map_SAm  + 
  transition_manual(Date)

# 6. Creating gif

SAm_Gif <-  animate(Map_SAm_Anim, fps = 5, width = 950, height = 750)

# 7. Saving the gif created (Maybe should not insert)

anim_save("7.SAm.gif", animation =SAm_Gif, path = "C:/Users/Gustav/Desktop/MSc Data Science 1st year/Exploratory data analysis/Assignments/Assignment4" )

```


```{r Importing the merged gif file}

# Doesnt seem to work

Covid_19_gif <- read.gif("Ass_4_gif.gif")

```

```{r Test with faceting - Europe}

# Steps:

# 1. Filtering the spatial data frame to only retrieve information about Europe

ncov_new_Spatial_Eur <- ncov_new_Spatial %>% 
  filter(continent == "Europe")

# 2. Specifying the position of the labels for each country in Africa (counting the number of cases as the date progresses)

World_Points_Eur <- st_centroid(ncov_new_Spatial_Eur)
World_Points_Eur <- cbind(ncov_new_Spatial_Eur,
                          st_coordinates(st_centroid(ncov_new_Spatial_Eur$geometry)))

# 3. Getting data for base map (Europe countries)

Europe_data <- World %>% 
  filter(continent == "Europe")

# 4. Plotting Europe map using ggplot() 

Map_Eur <- ggplot() + 
  geom_sf(data = Europe_data, fill = "gray48") +
  geom_sf(data = ncov_new_Spatial_Eur, aes(frame = Date, fill = Cum_Confirmed_Country)) +
  coord_sf(xlim = c(-40, 100), ylim = c(30,75), expand = FALSE) +
  scale_fill_gradient(low="white", high="orange", trans = "log10") +
  geom_text(data = World_Points_Eur, aes(x = X, y = Y, label = Cum_Confirmed_Country), check_overlap = FALSE) + 
  labs(title = "Confirmed cases", subtitle = "Date: {current_frame}", fill = "Count", x = "Latitude", y = "Longitude" ) + 
  theme(title = element_text(size = 8, face = "bold"), panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))


Map_Eur_Deaths <- ggplot() + 
  geom_sf(data = Europe_data, fill = "gray48") +
  geom_sf(data = ncov_new_Spatial_Eur, aes(frame = Date, fill = Deaths_Country)) +
  coord_sf(xlim = c(-40, 100), ylim = c(30,75), expand = FALSE) +
  scale_fill_gradient(low="white", high="red", trans = "log10") +
  geom_text(data = World_Points_Eur, aes(x = X, y = Y, label = Deaths_Country), check_overlap = FALSE) + 
  labs(title = "Deaths", subtitle = "Date: {current_frame}", fill = "Count", x = "Latitude", y = "Longitude" ) + 
  theme(title = element_text(size = 8, face = "bold"), panel.grid.major = element_line(color = "gray65", linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))

Plot_Test <- plot_grid(Map_Eur, Map_Eur_Deaths)

Plot_Test_Anim <-  Plot_Test + transition_manual(Date)

animate(Plot_Test_Anim, fps = 5, width = 950, height = 750)



```


Part 2: Shiny app development

```{r Shiny app development}

# Steps:

# 1. Saving variables that are going to be used in the shiny application

Death_icon <- awesomeIcons(
  icon = "times-circle",
  library = "fa",
  iconColor = "black",
  markerColor = "red"
)

Recovery_icon <- awesomeIcons(
  icon = "ambulance",
  library = "fa",
  iconColor = "black",
  markerColor = "green"
)

# Adding the continent variavble to data frame

World_Merge_No_Geom <- World_Merge_df %>% 
  st_set_geometry(value = NULL)


Continents_data <- left_join(ncov_new_FINAL_data, World_Merge_No_Geom, by = "Country_Region") %>%
  group_by(Date, continent) %>% 
      summarise(Cum_Confirmed_Country = sum(Cum_Confirmed_Country),
            Recovered_Country = sum(Recovered_Country),
            Deaths_Country = sum(Deaths_Country),
            Infected_Country = sum(Infected_Country)) %>% 
  rename(Country_Region = continent) %>% 
  as.data.frame()


ncov_new_FINAL_data_Plot <- rbind(Continents_data, ncov_new_FINAL_data)



# 1.1 Country names and continetn names


Country_Names <- levels(as.factor(ncov_new_FINAL_data$Country_Region))

Continent_Names <- levels(as.factor(Continents_data$Country_Region))

Country_Region_Names <- c(Continent_Names, Country_Names)  


# 1.2 Creating Ranking table of countries

Rank_Country_Table <- ncov_new_FINAL_data %>% 
       subset(Date == "2020-03-23") %>% 
       select(Country_Region, Cum_Confirmed_Country) %>% 
     arrange(desc(Cum_Confirmed_Country)) %>%   
      rename(Country = Country_Region,
            Cases = Cum_Confirmed_Country )

# 1.3 Filtering to the desired date of user to reveal map and creating colour pallet

Map_Data <- ncov_new_Spatial %>% 
  filter(Date =="2020-03-23")

Countries_pal <- colorNumeric(palette = "Reds",
                              domain = log(Map_Data$Cum_Confirmed_Country))

# 1.4 Adding reactive variable

```
 
Begin app here!
```{r Actual beginning of app}

#2.1 Define all the elements of the  header, sidebar and body of the dashboard. 

header <- dashboardHeader(
  title = "Covid 19 - 23/03/2020",
  titleWidth = 250
  
)

sidebar <- dashboardSidebar(
  
# Setting the deafault of the sidebar to close  
collapsed = TRUE  
# Adding tabs to the sidebar in the dashboard
 ,  
sidebarMenu(
menuItem("Word Data",
tabName = "world_data", icon = icon("globe")),
menuItem("Country/Region Data",
tabName = "country_region_data", icon = icon("chart-bar"))
 )
)
  
body <- dashboardBody(
  
  # tags$head(tags$style(HTML('
  #     .main-header .logo {
  #       font-family: "Courier New";
  #       font-weight: bold;
  #       font-size: 24px;
  #     }
  #   '))),
  
# Linking the tabs created in the sidebar to reflect in the body of the dashboard
  
 tabItems(
tabItem(tabName = "world_data",

fluidRow(
  
  column( width = 3,
          valueBoxOutput("G_Confirmed_Cases",
                         width = NULL
          )
          
  ),
  column( width = 3,
          valueBoxOutput("G_deaths",
                         width = NULL)
          
  ),
  column(width = 3,
         valueBoxOutput("G_Recoveries",
                        width = NULL)
  ),
  
  column(width = 3,
         valueBoxOutput("G_Infected",
                        width = NULL)
   )
  
  
),

fluidRow(

# Creating space for leaflet and table output 
  
leafletOutput("World_Map", height = 550)
      
 #  column(width = 3,
 #  dataTableOutput("World_Table"))
 # 

  ),

fluidRow(
  column(width = 6,
          align = "center",
  setSliderColor("red", 1),    
  sliderInput("date","Select data extraction date:",
              min = as.Date("2020-01-22","%Y-%m-%d"),
              max = as.Date("2020-03-23","%Y-%m-%d"),
              value=as.Date("2020-03-23"),
              timeFormat="%Y-%m-%d")
            
  ),

  column(width = 6,
          align = "left",
         prettyRadioButtons(inputId = "topTen", 
                            label = "Additional information w.r.t chosen date:",
                            choices = c("Normal world map", "Top ten regions with highest death count", "Top ten regions with the higest recovery count"),
                          shape = "round", 
                          status = "danger",
                          fill = TRUE, 
                          animation = "pulse")
                     
                     )  

 )
),

tabItem(tabName = "country_region_data",
        
        sidebarLayout(
          sidebarPanel(
 
                       
            selectInput(inputId = "country_continent",
                        "Select Country/Continent:",
                        choices = Country_Region_Names ,
                        selected = "South Africa", 
                        multiple = TRUE),
            
            prettyRadioButtons(inputId = "DataSelection",
                          label = "Select data interest:", 
                          icon = icon("check-circle"),
                          choiceNames = c("Cumulative cases", "Deaths", "Recoveries", "Infected"),
                          choiceValues = c("Cum_Confirmed_Country", "Deaths_Country", "Recovered_Country", "Infected_Country"),
                          selected = "Cum_Confirmed_Country",
                          animation = "tada", 
                          status = "danger"),
            
            materialSwitch(inputId = "logScale",
                           label = "Activate log scale:",
                           status = "danger"
                           ),
            
            materialSwitch(inputId = "area",
                           label = "Activate area graph:",
                           status = "danger"
                           )
            
          ),
          
          
          
          mainPanel(
            plotlyOutput("countryPlot"),
            
            plotlyOutput("FreqPlot")
          )
          
        )
        

        
   )
  )
 )


ui <- dashboardPage(header, sidebar, body, skin = "black")



server <- function(input, output){

# Steps:

 
  
# 1. Saving reactive variables to minimise duplication of code
  
# 1.1 Data set used for value box inputs  
ncov_reactive_data <-  reactive({
  ncov_new_FINAL_data %>% 
    filter(Date == input$date)
})

#


# 2. Specifying value box information

# 2.1 Value box 1 output  
output$G_Confirmed_Cases <- renderValueBox({
  
  Confirmed <-  ncov_reactive_data() %>% 
    select(Cum_Confirmed_Country) %>% 
    sum()
  
  valueBox(value = Confirmed,
           subtitle = "Global confirmed cases",
           icon = icon("globe", lib = "glyphicon"),
                       color = "red")
  })

# 2.2 Value box 2 output
output$G_deaths <- renderValueBox({
  
  Deaths <- ncov_reactive_data() %>% 
    select(Deaths_Country) %>% 
    sum()
  
  valueBox(value = Deaths,
           subtitle = "Global deaths",
           icon = icon("remove-sign", lib = "glyphicon"),
                       color = "red")
})

# 2.3 Value box 3 output

output$G_Recoveries <- renderValueBox({
  
  Recoveries <- ncov_reactive_data() %>% 
    select(Recovered_Country) %>% 
    sum()
  
  valueBox(value = Recoveries,
           subtitle = "Global recoveries",
           icon = icon("heart", lib = "glyphicon"),
                       color = "red")
})

# 2.4 Value box 4 output
output$G_Infected <- renderValueBox({
  
  Infected <- ncov_reactive_data() %>% 
    select(Infected_Country) %>% 
    sum()
  
  valueBox(value = Infected,
           subtitle = "Global infected cases",
           icon = icon("stethoscope", lib = "font-awesome"),
                       color = "red")
})

# 3. Output for world map



output$World_Map <- renderLeaflet({ 
  
  Top10Deaths <- ncov_new %>%
  filter(Date == input$date) %>% 
  arrange(desc(Deaths)) %>% 
  head(10) %>% 
  mutate(Rank = seq(1,10, by = 1))
  
  Top10Recoveries <- ncov_new %>%
  filter(Date == input$date) %>% 
  arrange(desc(Recovered)) %>% 
  head(10) %>% 
  mutate(Rank = seq(1,10, by = 1))
  
  
  
  L_World_Map <- ncov_new_Spatial %>%
    filter(Date == input$date) %>% 
     leaflet(options = leafletOptions(minZoom = 2.5)) %>% 
       addProviderTiles(provider = "CartoDB.Positron", options = providerTileOptions( noWrap = TRUE) ) %>%  
       addPolygons(weight = 1,
                   fillOpacity = 0.2,
                   color = ~Countries_pal(log(Cum_Confirmed_Country)),
                   popup = ~paste0("<b>", Country_Region, "</b>","<br/>",
                                   "Cases:", Cum_Confirmed_Country,"<br/>",
                                   "Deaths:", Deaths_Country, "<br/>",
                                   "Recoveries:", Recovered_Country,"<br/>",
                                   "Infected:", Infected_Country),
                   highlight = highlightOptions(weight = 3,
                                                color = "red",
                                                bringToFront = TRUE)
       ) %>% 
       setMaxBounds(lat1 = -90, lat2 = 90, lng1 = -180, lng2 = 180) %>% 
       setView(lat = 8.0817, lng = 17.6078, zoom = 2.5)
  
   if(input$topTen == "Normal world map") {
    
    L_World_Map
    
  }
  
  if(input$topTen == "Top ten regions with highest death count") {
  
    L_World_Map <- L_World_Map %>% 
      addAwesomeMarkers(lng = Top10Deaths$Long, lat = Top10Deaths$Lat, icon = Death_icon, popup = ~paste0("<b>","Rank:",Top10Deaths$Rank,"</b>","<br/>","<b>","Deaths:", "</b>",Top10Deaths$Deaths))
    
  }
  
  if(input$topTen == "Top ten regions with the higest recovery count") {
    
   L_World_Map <- L_World_Map %>% 
      addAwesomeMarkers(lng = Top10Recoveries$Long, lat = Top10Recoveries$Lat, icon = Recovery_icon, popup = ~paste0("<b>","Rank:", Top10Recoveries$Rank,"</b>", "<br/>","<b>","Recoveries:", "</b>",Top10Recoveries$Recovered))
    
  }
 
  
  L_World_Map
  

  })



# 4. Plotly output


output$countryPlot <- renderPlotly({

   # NB New variable intorduced ncov_final_data_plot (includes continent data)
   
  data <- ncov_new_FINAL_data_Plot %>% 
    filter(Country_Region %in% input$country_continent) 
  

if(input$DataSelection == "Cum_Confirmed_Country") {
  
data <-  data %>% 
  filter(Cum_Confirmed_Country != 0)

y <- data$Cum_Confirmed_Country
  
  Plot <- ggplot(data, aes(x = Date, y = Cum_Confirmed_Country , fill = Country_Region)) +
    geom_col(aes(text = paste0('</br> Cumalative cases: ', Cum_Confirmed_Country,'</br> Date: ', Date, '</br> Country: ', Country_Region)), position = "dodge") +
    labs(y = "Cummalative cases", fill = "Region") +
    theme_bw()
}
  
if(input$DataSelection == "Deaths_Country") {
  
  data <-  data %>% 
  filter(Deaths_Country != 0)
  
  y <- data$Deaths_Country
  
  Plot <- ggplot(data, aes(x = Date, y = Deaths_Country, fill = Country_Region)) +
    geom_col(aes(text = paste0('</br> Deaths: ', Deaths_Country,'</br> Date: ', Date, '</br> Country: ', Country_Region)), position = "dodge") +
    labs(y = "Deaths", fill = "Region") +
    theme_bw()
}

if(input$DataSelection == "Recovered_Country") {
  
  data <-  data %>% 
  filter(Recovered_Country != 0)
  
  y <- data$Recovered_Country
  
  Plot <- ggplot(data, aes(x = Date, y = Recovered_Country, fill = Country_Region)) +
    geom_col(aes(text = paste0('</br> Recoveries: ', Recovered_Country,'</br> Date: ', Date, '</br> Country: ', Country_Region)), position = "dodge") +
    labs(y = "Recoveries", fill = "Region") +
    theme_bw()
    
}
  
if(input$DataSelection == "Infected_Country") {
  
  data <-  data %>% 
  filter(Infected_Country != 0)
  
  
  y <- data$Infected_Country
  
  Plot <- ggplot(data, aes(x = Date, y = Infected_Country, fill = Country_Region)) +
    geom_col(aes(text = paste0('</br> Infected: ',Infected_Country,'</br> Date: ', Date, '</br> Country: ', Country_Region)), position = "dodge") +
    labs(y = "Infected", fill = "Region") +
    theme_bw()
}

  if(input$logScale){
    
  Plot <-  Plot + scale_y_log10() 
    
  }
  
  if(input$area){

Plot <- Plot %>% 
  remove_geom(geom = "col")

Plot <- Plot + geom_area(alpha = 0.6, 
                         position = "identity",
                         aes(fill = fct_reorder(Country_Region, y, .desc = TRUE)))

  }
  
if(input$area)  {
    
  ggplotly(Plot, tooltip = c("y","x")) %>% 
    layout(hovermode = "compare")
}
  else {
    ggplotly(Plot, tooltip = "text") %>% 
    layout(hovermode = "compare")
    
  }
})

 

}

shinyApp(ui = ui, server = server )


# thing to add
# 4. addResetMapButton() to reset teh map to its default position
   
# World table output
   # output$World_Table <- renderDataTable({Rank_Country_Table})
   

```


```{r ggplot example}

 dataTest <- ncov_new_FINAL_data %>% 
  filter(Country_Region == "China" | Country_Region == "Italy")


p <- ggplot(dataTest, aes(x = Date, y = Cum_Confirmed_Country, fill = Country_Region)) +
  geom_col(aes(text = paste0('</br> Cumalative count: ', Cum_Confirmed_Country,'</br> Date: ', Date, '</br> Country: ', Country_Region)), position = "dodge") 
  

# Density does not work but the one above does
y <- dataTest$Cum_Confirmed_Country

q <-  ggplot(dataTest, aes(x = Date, y = Cum_Confirmed_Country, fill = Country_Region)) +
  geom_area(aes(fill= fct_reorder(Country_Region,y, .desc = TRUE)), alpha = 0.7)
  

 ggplotly(q, tooltip = c("x", "y"))
 
 fig + ggplotly()
 
 
 tooltip = "text"
 
 
  ggplotly(q, tooltip = "label")
    
,)
 aes() 
 
 plot_ly(data = dataTest,
         x = ~Date,
         y = ~Cum_Confirmed_Country,
         type = "scatter",
         mode = "lines",
         fill = ~Country_Region)
 
 
 
  

```

```{r}
ui <- fluidPage(

    tags$h1("Modal with spinner"),
    actionButton("sleep1", "Launch a long calculation"),
    actionButton("sleep2", "And another one")
  )

  server <- function(input, output, session) {

    observeEvent(input$sleep1, {
      show_modal_spinner()
      Sys.sleep(5)
      remove_modal_spinner()
    })

    observeEvent(input$sleep2, {
      show_modal_spinner(
        spin = "cube",
        color = "red",
        text = "Please wait..."
      )
      Sys.sleep(5)
      remove_modal_spinner()
    })

  }

  shinyApp(ui, server)




```


**** Example ****


As seen by figure \ref{fig:Normal_Distribution}, the normal distribution is plotted.
Remember the default figure width height is 6 x 4.5 

```{r Normal_Distribution, fig.height= 3, fig.width=4.5 , fig.align='center', fig.cap= "The normal distribution curve", echo=FALSE}

x <- rnorm(100)

hist(x)

ncov_new_FINAL_data %>%
  filter(Date == "2020-03-23") %>% 
  select(Cum_Confirmed_Country) %>%
    sum()
  


```


\pagebreak

\section{References}

