---
title: "Assignment 4"
author: "Gustav Oosthuizen - OSTGUS001"
date: '`r format(Sys.Date(), "%d-%B-%Y")`'
output:
  pdf_document:
    highlight: pygments
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_caption: true
citation_package: "biblatex"
bibliography: "Ass_4.bib"
    
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\pagebreak

```{r installing packages, include=FALSE}

#Steps:

# 1. Installing "basic" packages:

#install.packages("tidyverse")
#install.packages("remotes")

library(tidyverse)
library(remotes)

# 2. Installing packages used to get corona virus data set

#remotes::install_github('pzhaonet/ncovr')
#install.packages("ncovr")
#library(ncovr)

```

```{r Loadinging and transforming most current data, include=FALSE}

# # CODE DOES NOT WORK
# 
# # Note: Only run the following chunk if the latest conona data wants to be obtained:
# 
# # Run the following command to get the latest data
# ncov <- get_ncov()
# 
# # To only extract the "area" data frame
# ncov <- ncov$area
# 
# # a function for converting Julian time to ymd_
# conv_time <- function(x){
#   as.POSIXct('1970-01-01', tz = 'GMT') + x / 1000
# }
# 
# # convert updateTime from Julian date to ymd-hms
# ncov$updateTime <- conv_time(ncov$updateTime)
# 
# 
# # Will save the latest data 
# save(ncov, file = "ncov-latest.Rdata")

```

\section{Data wrangling} \label{DW section}

The data wrangling consist of importing, tidying and transforming the data set [@wickham2016r].

\subsection{Imporing data} 

As mentioned in section \ref{DW section} the first step of data wrangling is importing the data.
```{r importing data, echo=FALSE}

# Importing the data set "ncov"

load("C:/Users/Gustav/Desktop/MSc Data Science 1st year/Exploratory data analysis/Assignments/Assignment4/ncov-latest (1).Rdata")

head(ncov, 5)

```  

\section{Visualisations}

As seen by figure \ref{fig:Normal_Distribution}, the normal distribution is plotted.
Remember the default figure width height is 6 x 4.5 

```{r Normal_Distribution, fig.height= 3, fig.width=4.5 , fig.align='center', fig.cap= "The normal distribution curve", echo=FALSE}

x <- rnorm(100)

hist(x)


```

\pagebreak

\section{References}


